---
title: "FigureS7_Promoter_responsiveness_singleEnhancers_MMA20240530"
author: "MMA"
date: "2023-08-05"
output: html_document
---

```{r knitr setup, message=FALSE, warning=FALSE,include=FALSE}
knitr::opts_chunk$set(
                      fig.path = "/DATA/usr/m.martinez.ara/GitHub/EEPCombinations/Rproj/Revision",
                      fig.align = 'center', 
                      fig.width = 5, 
                      fig.height = 5,
                      fig.pos = "H",
                      message = FALSE,
                      warning = FALSE,
                      dev = c("pdf"),
                      dpi = 500,
                      include = TRUE)


```

```{r libs and fun, include = FALSE, tidy = FALSE}
library(tidyverse)
library(here)
library(magrittr)
library(reshape2)
library(knitr)
library(gridExtra)
library(pander)
library(cowplot)
library(circlize)
library(ggplot2)
library(GGally)
library(scales)
library(heatmaply)
library(ggpubr)
library(ggrastr)
```

--- 

# Intro

Here we quantify promoter responsiveness to single enhancers relative to the average boost index across the 8 promoters.  
These plots are included in supplementary figure 7 and are a replica of the plots in Figure 5 but using single enhancers instead of pairs of enhancers.

# Load data

```{r}
load('/DATA/usr/m.martinez.ara/GitHub/EEPCombinations/data/MMA20230805_EEP_Norm.RData')
```

## merge all libs

```{r}

tib_Split_EEP_All_normDual<-bind_rows(tib_Split_EEP_Sox2_normDual, tib_Split_EEP_Klf2_normDual, tib_Split_EEP_Otx2_normDual, tib_Split_EEP_FGF5_normDual, tib_Split_EEP_Lefty1_normDual, tib_Split_EEP_Nanog_normDual, tib_Split_EEP_P14_normDual, tib_Split_EEP_Tbx3_normDual)
```

## PrepData

```{r}
tib_Split_EEP_All_normDual_model<-tib_Split_EEP_All_normDual%>%
  filter(!activity_all%in%c(NaN, Inf, -Inf, 0))%>%
  mutate(Type=ifelse(suffix1=="Controls" & suffix2=="Controls", "Controls", ifelse(
    suffix1=="Controls" & !suffix2=="Controls", "Singlets", ifelse(
    !suffix1=="Controls" & suffix2=="Controls", "Singlets", "Duplets" ))))%>%
      mutate(frag=ifelse(Type=="Controls", "Controls", ifelse(
        Type=="Singlets" & suffix2=="Controls", frag1, ifelse( 
          Type=="Singlets" & suffix1=="Controls" , frag2, "Duplets"
        ))))%>%
  mutate(linboost=activity_all/Baseline)
```

# Calculate average enhancer effects

```{r}
#transform to wider dataset
PairsObjectCorr <-tib_Split_EEP_All_normDual %>%
    filter(!activity_all%in%c(NaN, Inf, -Inf, 0))%>%
  mutate(Type=ifelse(suffix1=="Controls" & suffix2=="Controls", "Controls", ifelse(
    suffix1=="Controls" & !suffix2=="Controls", "Singlets", ifelse(
    !suffix1=="Controls" & suffix2=="Controls", "Singlets", "Duplets" ))))%>%
      mutate(frag=ifelse(Type=="Controls", "Controls", ifelse(
        Type=="Singlets" & suffix2=="Controls", frag1, ifelse( 
          Type=="Singlets" & suffix1=="Controls" , frag2, "Duplets"
        ))))%>%
  filter(Type=="Singlets")%>%
  filter(!suffix1=="Controls")%>%
  select(frag1, Promoter, boost)%>%
  group_by(frag1, Promoter)%>%
  summarise(boost=mean(boost))%>%
  ungroup()%>%
  pivot_wider(names_from = Promoter, values_from = boost)

PairsObjectCorr2 <-tib_Split_EEP_All_normDual %>%
    filter(!activity_all%in%c(NaN, Inf, -Inf, 0))%>%
  mutate(Type=ifelse(suffix1=="Controls" & suffix2=="Controls", "Controls", ifelse(
    suffix1=="Controls" & !suffix2=="Controls", "Singlets", ifelse(
    !suffix1=="Controls" & suffix2=="Controls", "Singlets", "Duplets" ))))%>%
      mutate(frag=ifelse(Type=="Controls", "Controls", ifelse(
        Type=="Singlets" & suffix2=="Controls", frag1, ifelse( 
          Type=="Singlets" & suffix1=="Controls" , frag2, "Duplets"
        ))))%>%
  filter(Type=="Singlets")%>%
  filter(!suffix2=="Controls")%>%
  select(frag2, Promoter, boost)%>%
  group_by(frag2, Promoter)%>%
  summarise(boost=mean(boost))%>%
  ungroup()%>%
  pivot_wider(names_from = Promoter, values_from = boost)
```


```{r}
#calculate avg effects
AvgEffects<-PairsObjectCorr%>%
                drop_na()%>%
mutate(AvgEff=(P7_Ap1m1+P3_OTX2+P2_KLF2+P6_Nanog+P1_SOX2+P4_FGF5+P8_Tbx3+P5_Lefty1)/8)%>%
  select(frag1, AvgEff)

AvgEffects2<-PairsObjectCorr2%>%
                drop_na()%>%
mutate(AvgEff=(P7_Ap1m1+P3_OTX2+P2_KLF2+P6_Nanog+P1_SOX2+P4_FGF5+P8_Tbx3+P5_Lefty1)/8)%>%
  select(frag2, AvgEff)
```

```{r}
# join datasets

tib_Split_EEP_All_normDual_PlusEffectsAvg<-tib_Split_EEP_All_normDual_model%>%
  #filter(!Promoter=="P7_Ap1m1")%>%
  filter(Type=="Singlets")%>%
    filter(!suffix1=="Controls")%>%
  left_join(AvgEffects, by="frag1")%>%
  filter(!boost%in%c(NA, Inf, -Inf, NaN))%>%
  filter(!AvgEff%in%c(NA, Inf, -Inf, NaN))

tib_Split_EEP_All_normDual_PlusEffectsAvg2<-tib_Split_EEP_All_normDual_model%>%
  #filter(!Promoter=="P7_Ap1m1")%>%
  filter(Type=="Singlets")%>%
    filter(!suffix2=="Controls")%>%
  left_join(AvgEffects2, by="frag2")%>%
  filter(!boost%in%c(NA, Inf, -Inf, NaN))%>%
  filter(!AvgEff%in%c(NA, Inf, -Inf, NaN))

tib_Split_EEP_All_normDual_PlusEffectsAvgComb<-
  bind_rows(tib_Split_EEP_All_normDual_PlusEffectsAvg,tib_Split_EEP_All_normDual_PlusEffectsAvg2)

```

```{r CorrBoostAvgEff_SingleEs, fig.height=8, fig.width=8}
# plot avg effect vs effect per promoter

tib_Split_EEP_All_normDual_PlusEffectsAvgComb%>%
   ggplot(aes(AvgEff, boost))+
  geom_point(alpha=0.25, size=0.4)+
  coord_fixed(0.5)+
  theme_bw()+
  geom_smooth(method = "lm")+
  geom_abline(slope = 1, color="grey")+
  stat_cor(
  label.y = -1.5)+
  facet_wrap(~Promoter)+
  xlab("Boost index (log2)")+
  xlab("Average Boost index (log2)")


```

# extract slopes

```{r}

# fit model to extract promoter specific slopes


fit1_AvgEff <- lm(boost ~ AvgEff + AvgEff:Promoter, data=tib_Split_EEP_All_normDual_PlusEffectsAvgComb)
summary(fit1_AvgEff)
  
```

```{r Slopes_AVGeff_singleEs, fig.width=5, fig.height=5}
# plot slopes vs baselines 

#extract baselines
Baselines<-tib_Split_EEP_All_normDual%>%
  #filter(Promoter=="P2_KLF2")%>%
  filter(!Additivebaseline%in%c(NA, Inf,-Inf))%>%
 #mutate(Supradd=ifelse(activity_all> (Additivebaseline+AdditiveSd) ,1, 0))%>%
  group_by(Promoter, Baseline)%>%
  summarise(N=n())

#extract slopes
            fit1_AvgEff$coefficients
 SlopesAvgEff<- c(1.404365, 1.404365-0.-0.805332, 1.404365-0.789060, 1.404365-0.075744, 1.404365+0.153797, 1.404365-0.697107, 1.404365-1.175281, 1.404365+0.027254 )    
 
 #plot
bind_cols(Baselines, slope=SlopesAvgEff)%>%
 ggplot(aes( log2(Baseline), slope))+
  geom_point(size=1.5, alpha=0.1)+
    #coord_fixed()+
  theme_bw()+
  geom_smooth(method = "lm", alpha=0.5)+
  #geom_abline(slope = 1)+
  stat_cor()+
    geom_point(aes(color=Promoter), size=3)+
xlab("log2(Promoter Activity)")
  
```

