---
title: "Figure2_SingleEnhancers_MMA20230805"
author: "MMA"
date: '2023-08-05'
output: html_document
---
```{r knitr setup, message=FALSE, warning=FALSE,include=FALSE}
knitr::opts_chunk$set(
                      fig.path = "/DATA/usr/m.martinez.ara/GitHub/EEPCombinations/Rproj/",
                      fig.align = 'center', 
                      fig.width = 5, 
                      fig.height = 5,
                      fig.pos = "H",
                      message = FALSE,
                      warning = FALSE,
                      dev = c("pdf"),
                      dpi = 500,
                      include = TRUE)


```

```{r libs and fun, include = FALSE, tidy = FALSE}
library(tidyverse)
library(here)
library(magrittr)
library(reshape2)
library(knitr)
library(gridExtra)
library(pander)
library(cowplot)
library(circlize)
library(datteRo)
library(ggplot2)
library(GGally)
library(scales)
library(heatmaply)
library(ggpubr)
library(ggrastr)
```

---
# Intro
In this document we analyze the effects of single enhacers across enhancer-control combinations and generate publication grade plots for Figure 2 of Martinez-Ara et al. 2023.

# Funcs

```{r ploot matrix function, include=FALSE}

#Function to plot the Boosting index Matrix 
plot_coop_heatmap_EEP <- function(tib, classs = 'all' , skip_n_lab = 1){

  if(classs != 'all') {
    # note: tib %>% filter(class == get('class')) fails
    # see https://stackoverflow.com/questions/40169949/filter-dataframe-using-global-variable-with-the-same-name-as-column-name

    tib     <- tib %>% filter(class %in% class)
  }
  # set color palette
  pal <- c('navyblue', 'gray95', 'orangered')

  # static viz
  print("1")

      # compute coop index range
      y_min <- min(tib[['boost']], na.rm = TRUE)
      y_max <- max(tib[['boost']], na.rm = TRUE)

      # coop index heatmap (additive)
      p_main <- tib %>%
        ggplot(aes(x = id1, y = id2)) +
        geom_tile(fill = 'gray40') +
        #        geom_tile(color = 'gray20', alpha = 0, size = 0.5) + ## prev version with stroke
        #        geom_tile(fill = rgb(53/255, 143/255, 168/255, 0.1)) + ## prev version with color outside coop idx palette
        geom_raster(data = tib, aes(x = id1, y = id2, fill = boost))
    print("2")
    # cosmetics
    p_main <- p_main +
      scale_fill_gradientn(colors = pal,
                           values = scales::rescale(c(y_min, 0, y_max))) +
      labs(x = 'Fragment 1', y = 'Fragment 2', fill = 'Boost Index', caption = paste0('(n = ', nrow(tib), ' pairs)')) +
      theme(axis.text.x      = element_text(angle = 90, color =  c('black', rep('transparent', skip_n_lab))),
            axis.text.y      = element_text(color =  c('black', rep('transparent', 0))),
            panel.grid       = element_blank(),
            panel.background = element_rect(fill = "white"),
            legend.position  = 'bottom',
            text             = element_text(size = 13),)+
      ggtitle((tib$Promoter)[[1]])

    print("3")

  p_main



}
```

# Load data

```{r}
load('/DATA/usr/m.martinez.ara/GitHub/EEPCombinations/data/MMA20230805_EEP_Norm.RData')
```

## merge all libs

```{r}

tib_Split_EEP_All_normDual<-bind_rows(tib_Split_EEP_Sox2_normDual, tib_Split_EEP_Klf2_normDual, tib_Split_EEP_Otx2_normDual, tib_Split_EEP_FGF5_normDual, tib_Split_EEP_Lefty1_normDual, tib_Split_EEP_Nanog_normDual, tib_Split_EEP_P14_normDual, tib_Split_EEP_Tbx3_normDual)
```



# Aim

Here I want to quantify the nr of significantly active enhancers there is per promoter and across promoters. 


# Analysis

## per promoter

```{r fig.height=3, fig.width=3.5}
#prepare data
tib_Split_EEP_All_normDual_wilcox <- tib_Split_EEP_All_normDual%>%
  mutate(Type=ifelse(suffix1=="Controls" & suffix2=="Controls", "Controls", ifelse(
    suffix1=="Controls" & !suffix2=="Controls", "Singlets", ifelse(
    !suffix1=="Controls" & suffix2=="Controls", "Singlets", "Duplets" ))))%>%
      mutate(frag=ifelse(Type=="Controls", "Controls", ifelse(
        Type=="Singlets" & suffix2=="Controls", frag1, ifelse( 
          Type=="Singlets" & suffix1=="Controls" , frag2, "Duplets"
        ))))
tib_Split_EEP_All_normDual_wilcox%>%
filter(frag=="Controls")

## For loop for wilcoxon test
WilcoxTestEnhancersPerP<-tibble(frag=NA, Promoter=NA, pvalue=NA, statistic=NA, n=NA , MedBoost=NA)


for (j in unique(tib_Split_EEP_All_normDual_wilcox$Promoter)){
for (i in unique(tib_Split_EEP_All_normDual_wilcox$frag)) {
  wicoxFtestE<-tib_Split_EEP_All_normDual_wilcox%>%
    ungroup()%>%
        filter(Type=="Singlets")%>%
    filter(frag==i)%>%
    filter(Promoter==j)%>%
    select(boost)
  wicoxFtestrest<-tib_Split_EEP_All_normDual_wilcox%>%
    ungroup()%>%
    filter(Type=="Controls")%>%
    filter(Promoter==j)%>%
    select(boost)
  if (nrow(wicoxFtestE)>5){
  WilcoxTestResult<-wilcox.test(wicoxFtestrest$boost,wicoxFtestE$boost)
  
  WilcoxTestEnhancersPerP<-WilcoxTestEnhancersPerP%>%
    add_row(frag=i,Promoter=j, pvalue=WilcoxTestResult$p.value, statistic=WilcoxTestResult$statistic,
            n=nrow(wicoxFtestE), MedBoost=median(wicoxFtestE$boost))
    }
  }
}
  
```


```{r}
WilcoxTestEnhancersPerP<-WilcoxTestEnhancersPerP%>%
  mutate(pAdj=p.adjust(pvalue))
```


```{r SignificantSingleEs_per_promoter}
WilcoxTestEnhancersPerP%>%
  filter(!is.na(MedBoost))%>%
  ggplot(aes(Promoter, MedBoost, fill=pAdj<(0.01)))+
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth =0.1)+
  theme_bw()
```


## Across promoters


```{r}
WilcoxTestEnhancersAcrossP<-tibble(frag=NA, pvalue=NA, statistic=NA, n=NA , MedBoost=NA)



for (i in unique(tib_Split_EEP_All_normDual_wilcox$frag)) {
  wicoxFtestE<-tib_Split_EEP_All_normDual_wilcox%>%
    ungroup()%>%
        filter(Type=="Singlets")%>%
    filter(frag==i)%>%
    select(boost)
  wicoxFtestrest<-tib_Split_EEP_All_normDual_wilcox%>%
    ungroup()%>%
    filter(Type=="Controls")%>%
    select(boost)
  if (nrow(wicoxFtestE)>5){
  WilcoxTestResult<-wilcox.test(wicoxFtestrest$boost,wicoxFtestE$boost)
  
  WilcoxTestEnhancersAcrossP<-WilcoxTestEnhancersAcrossP%>%
    add_row(frag=i, pvalue=WilcoxTestResult$p.value, statistic=WilcoxTestResult$statistic,
            n=nrow(wicoxFtestE), MedBoost=median(wicoxFtestE$boost))
    }
  }

```

```{r}
WilcoxTestEnhancersAcrossP<-WilcoxTestEnhancersAcrossP%>%
  mutate(pAdj=p.adjust(pvalue))
```


```{r SignificantSingleEs_across_promoters}
WilcoxTestEnhancersAcrossP%>%
  mutate(Promoter="All_Promoters")%>%
  ggplot(aes(Promoter, MedBoost, fill=pAdj<(0.01)))+
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth =0.1)+
  theme_bw()
```

## Plot
```{r eval=FALSE, fig.height=10, fig.width=10, include=FALSE}
# plot distributions per fragment per promoter

tib_Split_EEP_All_normDual_wilcox%>%
  filter(!Type=="Duplets")%>%
  ggplot(aes(frag, boost))+
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth =0.0055)+
  theme_bw()+
  facet_wrap(~Promoter)
  
```

```{r SignificantSingleEs_Sox2 , fig.width=9, fig.height=4}
# plot distributions per fragment per promoter

tib_Split_EEP_All_normDual_wilcox%>%
  filter(!Type=="Duplets")%>%
  filter(Promoter=="P1_SOX2")%>%
  filter(!boost%in% c(NA,Inf, -Inf))%>%
  left_join(WilcoxTestEnhancersPerP, by = c("Promoter", "frag"))%>%
  mutate(n=ifelse(n%in%c(NA) & Type=="Controls", 100, n))%>%
  filter(!is.na(n))%>%
  ggplot(aes(frag, boost))+
   #ggbeeswarm::geom_beeswarm(priority='density',size=0.5)
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth =0.025)+
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth =0.025, aes( color=pAdj<0.01))+
  stat_summary(fun.y = median, width = 0.6, linetype = "solid", geom = "crossbar")+
  #geom_boxplot(alpha=0.7, aes( fill=pAdj<0.01), outlier.shape = NA)+
  theme_bw(base_size = 10)+
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  ggtitle("E+C activation Sox2 promoter")
  #facet_wrap(~Promoter)
  
```


# mini matrix
```{r mini_matrix, fig.width=15, fig.height=8}
tib_Split_EEP_All_normDual%>%
  mutate(Type=ifelse(suffix1=="Controls" & suffix2=="Controls", "Controls", ifelse(
    suffix1=="Controls" & !suffix2=="Controls", "Singlets", ifelse(
    !suffix1=="Controls" & suffix2=="Controls", "Singlets", "Duplets" ))))%>%
      mutate(frag=ifelse(Type=="Controls", "Controls", ifelse(
        Type=="Singlets" & suffix2=="Controls", frag1, ifelse( 
          Type=="Singlets" & suffix1=="Controls" , frag2, "Duplets"
        ))))%>%
filter(!Type=="Duplets")%>%
  filter(!boost%in% c(NA, Inf, -Inf))%>%
  group_by(Promoter, frag)%>%
  summarise(boost=median(boost), N=n())%>%
  filter(N>10)%>%
    mutate(id1=frag, id2=Promoter)%>%
  #filter(log2(activity_all_frag2)!=c(NA, Inf,-Inf))%>%
  #filter(log2(activity_all_frag1)!=c(NA, Inf,-Inf))%>%
plot_coop_heatmap_EEP( class = c('all'), skip_n_lab = 0)
```


# quantify Es with generally activating effects

```{r}
WilcoxTestEnhancersAcrossP%>%
  filter(!is.na(pvalue))%>%
  mutate(activator=ifelse(pAdj<0.01 & MedBoost>0.01,1,0))%>%
  summarise(n=n(), nAct=sum(activator) , PercAct=sum(activator)/n)

WilcoxTestEnhancersAcrossP%>%
  filter(pAdj>0.01)
  
```

```{r}
WilcoxTestEnhancersPerP%>%
  filter(!is.na(pvalue))%>%
  mutate(activator=ifelse(pAdj<0.01 & MedBoost>0.01,1,0))%>%
  group_by(frag)%>%
  summarise(n=n(), nAct=sum(activator) , PercAct=sum(activator)/n, medboost=mean(MedBoost))
  
WilcoxTestEnhancersPerP%>%
  filter(!is.na(pvalue))%>%
  mutate(activator=ifelse(pAdj<0.01 & MedBoost>0.01,1,0))%>%
  group_by(frag)%>%
  summarise(n=n(), nAct=sum(activator) , PercAct=sum(activator)/n)%>%
  mutate(Act=ifelse(nAct>0,1,0))%>%
  summarise(Ntot=n(), nActaccP=sum(Act))

```


# Example effects across promoters


```{r Nanog_E074_example, fig.width=5, fig.height=5}
# plot distributions per fragment per promoter

tib_Split_EEP_All_normDual_wilcox%>%
  filter(Type=="Singlets")%>%
  #filter(Promoter=="P7_Ap1m1")%>%
  filter(!boost%in% c(NA,Inf, -Inf))%>%
  left_join(WilcoxTestEnhancersPerP, by = c("Promoter", "frag"))%>%
  #mutate(n=ifelse(n%in%c(NA) & Type=="Controls", 100, n))%>%
  #filter(!is.na(n))%>%
  filter(frag=="Nanog_E074")%>%
  ggplot(aes(Promoter, boost))+
   #ggbeeswarm::geom_beeswarm(priority='density',size=0.5)
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth =0.04)+
stat_summary(fun.y = median, width = 0.6, linetype = "solid", geom = "crossbar")+
  theme_bw(base_size = 12)+
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  ggtitle("E+C activation Nanog_E073  per promoter")
  #facet_wrap(~Promoter)
  
```


