---
title: "Figure4_Supra-SubAdditivity_MMA20230805"
author: "MMA"
date: "2023-08-05"
output: html_document
---
```{r knitr setup, message=FALSE, warning=FALSE,include=FALSE}
knitr::opts_chunk$set(
                      fig.path = "/DATA/usr/m.martinez.ara/GitHub/EEPCombinations/Rproj/",
                      fig.align = 'center', 
                      fig.width = 5, 
                      fig.height = 5,
                      fig.pos = "H",
                      message = FALSE,
                      warning = FALSE,
                      dev = c("pdf"),
                      dpi = 500,
                      include = TRUE)


```

```{r libs and fun, include = FALSE, tidy = FALSE}
library(tidyverse)
library(here)
library(magrittr)
library(reshape2)
library(knitr)
library(gridExtra)
library(pander)
library(cowplot)
library(circlize)
library(datteRo)
library(ggplot2)
library(GGally)
library(scales)
library(heatmaply)
library(ggpubr)
library(ggrastr)
```

--- 

# Intro

Here we analyse the supra and sub-additive interactions of enhancer-enhancer combinations and generate plots for Figure 4.

# Load data

```{r}
load('/DATA/usr/m.martinez.ara/GitHub/EEPCombinations/data/MMA20230805_EEP_Norm.RData')
```

## merge all libs

```{r}

tib_Split_EEP_All_normDual<-bind_rows(tib_Split_EEP_Sox2_normDual, tib_Split_EEP_Klf2_normDual, tib_Split_EEP_Otx2_normDual, tib_Split_EEP_FGF5_normDual, tib_Split_EEP_Lefty1_normDual, tib_Split_EEP_Nanog_normDual, tib_Split_EEP_P14_normDual, tib_Split_EEP_Tbx3_normDual)
```

# synergies Analysis

## per promoter

```{r fig.height=3, fig.width=3.5}
#prepare data
tib_Split_EEP_All_normDual_wilcox <- tib_Split_EEP_All_normDual%>%
  mutate(Type=ifelse(suffix1=="Controls" & suffix2=="Controls", "Controls", ifelse(
    suffix1=="Controls" & !suffix2=="Controls", "Singlets", ifelse(
    !suffix1=="Controls" & suffix2=="Controls", "Singlets", "Duplets" ))))%>%
      mutate(frag=ifelse(Type=="Controls", "Controls", ifelse(
        Type=="Singlets" & suffix2=="Controls", frag1, ifelse( 
          Type=="Singlets" & suffix1=="Controls" , frag2, "Duplets"
        ))))
tib_Split_EEP_All_normDual_wilcox%>%
   mutate(Baseline= ifelse(Promoter=="P4_FGF5", Baseline-0.07, Baseline))%>%
  mutate(Additivebaseline=Baseline1+Baseline2-Baseline)%>%
  mutate(AdditiveCoop==log2(activity_all/Additivebaseline))%>%
filter(frag=="Controls")

## For loop for wilcoxon test
WilcoxTestEnhancersPerP<-tibble(frag=NA, Promoter=NA, pvalue=NA, statistic=NA, n=NA , MedBoost=NA)


for (j in unique(tib_Split_EEP_All_normDual_wilcox$Promoter)){
for (i in unique(tib_Split_EEP_All_normDual_wilcox$frag1)) {
  wicoxFtestE<-tib_Split_EEP_All_normDual_wilcox%>%
    filter(!AdditiveCoop%in% c(NA, Inf, -Inf))%>%
    ungroup()%>%
        filter(Type=="Duplets")%>%
    filter(frag1==i | frag2==i)%>%
    filter(Promoter==j)%>%
    select(AdditiveCoop)
  wicoxFtestrest<-tib_Split_EEP_All_normDual_wilcox%>%
    ungroup()%>%
    filter(Type=="Duplets")%>%
    #in this case I compare to the rest of the duplets, not to controls
        filter(!frag==i & !frag2==i)%>%
    filter(Promoter==j)%>%
    select(AdditiveCoop)
  if (nrow(wicoxFtestE)>5){
  WilcoxTestResult<-wilcox.test(wicoxFtestrest$AdditiveCoop,wicoxFtestE$AdditiveCoop)
  
  WilcoxTestEnhancersPerP<-WilcoxTestEnhancersPerP%>%
    add_row(frag=i,Promoter=j, pvalue=WilcoxTestResult$p.value, statistic=WilcoxTestResult$statistic,
            n=nrow(wicoxFtestE), MedBoost=median(wicoxFtestE$AdditiveCoop))
    }
  }
}
  
```


```{r}
WilcoxTestEnhancersPerP<-WilcoxTestEnhancersPerP%>%
  mutate(pAdj=p.adjust(pvalue))
```


```{r SignificantSynergies_per_promoter}
WilcoxTestEnhancersPerP%>%
  filter(!is.na(MedBoost))%>%
  ggplot(aes(Promoter, MedBoost))+
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth =0.02)+
  theme_bw()+
  ylab("Obs/Exp")+
  ggtitle("General synergies per Enhancer and Promoter")

```
# Supra/Sub-additivity per promoter

Here we quantify how many combinations deviate more than 1 propagated standard deviation from the additive estimation.

```{r fig.height=3, fig.width=3.5}
#prepare data
tib_Split_EEP_All_normDual_PropSd <- tib_Split_EEP_All_normDual%>%
  filter(!is.na(Additivebaseline))%>%
  #propagateSDs square root of sum of squares of individual sds
  mutate(MulltipBAseline=Baseline1*Baseline2/Baseline)%>%
  mutate(AdditiveSd=sqrt((StDevBaseline^2)+(StDevBaseline1^2)+(StDevBaseline2^2)))%>%
   mutate(MultSd=MulltipBAseline*sqrt(((StDevBaseline/Baseline)^2)+((StDevBaseline1/Baseline1)^2)+((StDevBaseline2/Baseline2)^2)))%>%
  #is measured activity within Additive estimate +-1sd
  #filter((Additivebaseline-AdditiveSd)>0)
  mutate(SignifAddDev=ifelse(activity_all> (Additivebaseline-AdditiveSd) & activity_all< (Additivebaseline+AdditiveSd),"No", "Yes"))%>%
  mutate(SignifAddDev2sd=ifelse(activity_all> (Additivebaseline-2*AdditiveSd) & activity_all< (Additivebaseline+2*AdditiveSd),"No", "Yes"))%>%
mutate(SignifMultDev=ifelse(activity_all> (MulltipBAseline-MultSd) & activity_all< (MulltipBAseline+MultSd),"No", "Yes"))%>%
  mutate(SignifMultDev2sd=ifelse(activity_all> (MulltipBAseline-2*MultSd) & activity_all< (MulltipBAseline+2*MultSd),"No", "Yes"))%>%
  mutate(SupraddNotMult=ifelse(activity_all> (Additivebaseline+AdditiveSd) & activity_all< (MulltipBAseline),"Yes", "No"))%>%
mutate(SupraddMult=ifelse(abs(log2(MulltipBAseline/Additivebaseline))> 0.5 & activity_all> (MulltipBAseline),"Yes", "No"))%>%
mutate(ClosertoMult=ifelse(abs(log2(MulltipBAseline/Additivebaseline))> 0.5 & abs(activity_add-Additivebaseline)>abs(activity_all-MulltipBAseline),"Yes", "No"))

```

```{r 1SDAwayDist, fig.width=8, fig.height=4}
# plot distributions per fragment per promoter

p<-tib_Split_EEP_All_normDual_PropSd%>%
  #filter(Promoter=="P1_SOX2")%>%
  filter(!Additivebaseline%in%c(NA, Inf,-Inf))%>%
  mutate(Addboost=log2(Additivebaseline/Baseline))%>%
 ggplot(aes( Promoter, AdditiveCoop))+
  scale_color_manual(values=c("grey50", "turquoise"))+
geom_dotplot(binaxis = "y", stackdir = "center", binwidth =0.015, alpha=0.3, aes(fill=SignifAddDev, color=SignifAddDev))+
stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
                 geom = "crossbar", width = 0.75)+
    #coord_fixed()+
  theme_bw()+
  theme( # remove the vertical grid lines
           panel.grid.major.x = element_blank() ,
           # explicitly set the horizontal lines (or they will disappear too)
           panel.grid.major.y = element_line( size=.1 ) 
    )+
 scale_y_continuous(breaks = seq(from = -4, to = 8, by = 2))
  #geom_smooth(method = "lm")+
  #geom_abline(slope = 1)
  #stat_cor()
library(ggrastr)

  rasterize(p, layers='Dotplot', dpi=600)

tib_Split_EEP_All_normDual_PropSd%>%
  filter(!Promoter=="P4_FGF5")%>%
  filter(!Additivebaseline%in%c(NA, Inf,-Inf))%>%
  mutate(Addboost=log2(Additivebaseline/Baseline))%>%
 ggplot(aes( Promoter, AdditiveCoop))+
geom_dotplot(binaxis = "y", stackdir = "center", binwidth =0.01, alpha=0.3, aes(fill=SignifAddDev, color=SignifAddDev))+
  geom_boxplot(alpha=0.7, outlier.shape = NA)+
    #coord_fixed()+
  theme_bw()

  #geom_smooth(method = "lm")+
  #geom_abline(slope = 1)
  #stat_cor()
  
```

# Supra/Sub-additivity vs baseline

Here we plot the relationships between promoter baselines and Supra/sub-additivity

```{r SupraddVSBAseline, fig.width=5, fig.height=5}
# plot distributions per fragment per promoter

tib_Split_EEP_All_normDual_PropSd%>%
  #filter(Promoter=="P2_KLF2")%>%
  filter(!Additivebaseline%in%c(NA, Inf,-Inf))%>%
 mutate(Supradd=ifelse(activity_all> (Additivebaseline+AdditiveSd) ,1, 0))%>%
  group_by(Promoter, Baseline)%>%
  summarise(N=n(), SupraddN=sum(Supradd), PercSupradd=SupraddN/N)%>%
 ggplot(aes( log2(Baseline), PercSupradd))+
    #coord_fixed()+
  theme_bw()+
  geom_smooth(method = "lm", alpha=0.3)+
    geom_point(size=1)+
  #geom_abline(slope = 1)+
  stat_cor()+
    geom_point(aes(color=Promoter), size=3.5)+
  ylim(c(-0.05, 0.55))
  
```

```{r SubaddVSBAseline, fig.width=5, fig.height=5}
# plot distributions per fragment per promoter

tib_Split_EEP_All_normDual_PropSd%>%
  #filter(Promoter=="P2_KLF2")%>%
  filter(!Additivebaseline%in%c(NA, Inf,-Inf))%>%
 mutate(Supradd=ifelse(activity_all< (Additivebaseline-AdditiveSd) ,1, 0))%>%
  group_by(Promoter, Baseline)%>%
  summarise(N=n(), SupraddN=sum(Supradd), ProportionSubadd=SupraddN/N)%>%
 ggplot(aes( log2(Baseline), ProportionSubadd))+
    #coord_fixed()+
  theme_bw()+
  geom_smooth(method = "lm", alpha=0.3)+
    geom_point(size=1)+
  #geom_abline(slope = 1)+
  stat_cor()+
    geom_point(aes(color=Promoter), size=3.5)+
  ylim(c(-0.05, 0.55))
  
```

# Supra/Sub-additivity per enhancer and promoter

```{r Synergies_per_enhancer, fig.width=12, fig.height=5}
WilcoxTestEnhancersPerP%>%
  filter(!is.na(MedBoost))%>%
  ggplot(aes(reorder(frag, MedBoost), MedBoost, color=Promoter))+
  geom_point(binaxis = "y", stackdir = "center", binwidth =0.02)+
    stat_summary(fun.y = mean, width = 0.5, linetype = "solid", colour="darkgrey", geom = "crossbar")+
  theme_bw()+
  ylab("log2(Obs/Exp)")+
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  #geom_segment(aes(x = frag-0.5, xend = frag+0.5, y = median(MedBoost), yend = median(MedBoost)))
  ggtitle("General supra-additivity per Enhancer and Promoter")

WilcoxTestEnhancersPerP%>%
  filter(!is.na(MedBoost))%>%
  filter(MedBoost>0.5)%>%
  select(frag)%>%
  distinct

WilcoxTestEnhancersPerP%>%
  filter(!is.na(MedBoost))%>%
  filter(MedBoost>0.5)%>%
  select(frag)%>%
  group_by(frag)%>%
  summarise(n=n())%>%
  filter(n>1)

WilcoxTestEnhancersPerP%>%
  filter(!is.na(MedBoost))%>%
  select(frag)%>%
  distinct

19/58
6/58
```

# Enhancer Examples

## Nodal_E160

```{r Synergies_Examples, fig.width=5, fig.height=8}
# plot distributions per fragment per promoter

tib_Split_EEP_All_normDual_wilcox%>%
   mutate(Baseline= ifelse(Promoter=="P4_FGF5", Baseline-0.07, Baseline))%>%
  mutate(Additivebaseline=Baseline1+Baseline2-Baseline)%>%
  mutate(AdditiveCoop==log2(activity_all/Additivebaseline))%>%
  filter(Type=="Duplets")%>%
  filter(!AdditiveCoop%in% c(NA,Inf, -Inf))%>%
  filter(frag1=="Nodal_E160" | frag2=="Nodal_E160")%>%
  ggplot(aes(Promoter, AdditiveCoop))+
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth =0.02, aes(fill=Promoter, color=Promoter))+
stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
                 geom = "crossbar", width = 0.75)+
  theme_bw(base_size = 10)+
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  ggtitle("E+E Synergies Nodal_E160 per promoter")+
  ylim(c(-2.5, 7.5))

tib_Split_EEP_All_normDual_wilcox%>%
    mutate(Baseline= ifelse(Promoter=="P4_FGF5", Baseline-0.07, Baseline))%>%
  mutate(Additivebaseline=Baseline1+Baseline2-Baseline)%>%
  mutate(AdditiveCoop==log2(activity_all/Additivebaseline))%>%
  filter(Type=="Duplets")%>%
  filter(!AdditiveCoop%in% c(NA,Inf, -Inf))%>%
  filter(frag1=="Sox2_E180" | frag2=="Sox2_E180")%>%
  ggplot(aes(Promoter, AdditiveCoop))+
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth =0.02, aes(fill=Promoter, color=Promoter))+
stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
                 geom = "crossbar", width = 0.75)+
  theme_bw(base_size = 10)+
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  ggtitle("E+E Synergies Sox2_E180 per promoter")+
  ylim(c(-2.5, 7.5))
  #facet_wrap(~Promoter)


  
```


# Synergies between clusters


```{r Within_between, fig.width=4, fig.height=4}
# plot distributions per fragment per promoter

tib_Split_EEP_All_normDual_wilcox%>%
  filter(Type=="Duplets")%>%
  mutate(Clust=ifelse(cluster=="Rest", "Between", "Within"))%>%
  filter(!AdditiveCoop%in% c(NA,Inf, -Inf))%>%
  ggplot(aes(Clust, AdditiveCoop))+
   #ggbeeswarm::geom_beeswarm(priority='density',size=0.5)
  geom_dotplot(binaxis = "y", stackdir = "center", binwidth =0.01)+
stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
                 geom = "crossbar", width = 0.3, color="grey")+
  theme_bw(base_size = 10)+
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  stat_compare_means(method="wilcox.test")+
  ggtitle("E+E Synergies within/between Clusters")

  
```
